---
title: "One site only"
author: "Jill Bolland"
date: "14/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code below reads all the data and gets it into a dataframe with Site name as a column

```{r}
library(readxl)
library(data.table)

filename = "C:/Users/j/Documents/Uni/ComSci 760/Weather Project/Data/Data for Giovanni from Martin.xlsx"
sheetnames = excel_sheets(path = filename)

# open the data from the sheet specified

# create function
read_excel_sheet <- function(sheet_name, path) {
  x <- read_excel(path = path, sheet = sheet_name)
}

beachData <- lapply(sheetnames, read_excel_sheet, filename)
names(beachData) <- sheetnames

# need to get into dataframe so code for graphs and subsetting works
allBeachesData <- as.data.frame(rbindlist(beachData))

# Get a vector of column names
# dim(x)[1] gives the number of data points in x  -dim outpust (rows col)
# apply that fuction to the lists in beacData to give a vector of numbers of rows for each site
beachrows = sapply(beachData, function(x) dim(x)[1])
# code replicates each sheetname as many times as in the beachrows vector
beachSiteCol = rep(sheetnames, beachrows)

# add the beach sites column to the dataframe
allBeachesData$site = beachSiteCol

#  can do same as above for site characteristics and wind direction


```



Get just one site data into a dataframe

Use the sheetnames[i]  where i is 1 to 5 "Clarks"      "Narrow Neck" "Judges Bay"  "Weymouth"    "Milford"   

```{r}
site = sheetnames[5]

siteData = allBeachesData[allBeachesData$site == site,]
head(siteData)
```

```{r}
# get a boxplot of all the variables   - excluding wind direction
# individual box plots


# set up plot area 3 by 3
par(mfrow=c(3, 3))
for (i in 2:8) {
  boxplot(siteData[,i], main= paste0(site," ",names(siteData)[i]))  
}


```

```{r}

# line plots
par(mfrow=c(3, 3))
for (i in 2:8) {
  plot(siteData[,i], type = 'l', main= paste0(site," ",names(siteData)[i]))  
}

```

```{r}

# Density plot
par(mfrow=c(3, 3))

for (i in 2:8) {
    d =density(siteData[,i], na.rm = TRUE)
    plot(d, type="n", main= paste0(site," ",names(siteData)[i]))
    polygon(d, col="red", border="gray")
}
```


```{r}
# Histograms 

par(mfrow=c(3, 3))

for (i in 2:8) {
    hist(siteData[,i], main= paste0(site," ",names(siteData)[i]), col = 'blue')
}

# put back to 1 by 1 charts
par(mfrow=c(1, 1))
```
```{r}
# Wind direction graphs
#require(devtools)
#install_github('davidcarslaw/openair')
library(openair)

par(mfrow=c(1, 2))
siteWindData = siteData[c(1,7,8)]
colnames(siteWindData) = c('date','wd','ws')
# basic wind frequency plot
polarFreq(siteWindData)
windRose(siteWindData)

```
```{r}
library(s20x)
# scatter plot of Entero & rain
pairs20x(siteData[2:6])

# scatter plot of Entero & Wind & solar
pairs20x(siteData[c(2,7:8)])

```
######   Add a class column to the data
```{r}
siteData$class = ifelse(siteData$Entero < 280,0,1)

```

 
```{r}
# get the last 10% of data and keep as the test data
# we need to make sure the data is sorted in time and then calculate the number of observations and split the last 10 %

siteData = siteData[order(siteData$DATE), ]

# Drop the Entero level so we just have the class data  - also drop the site name column
#  Drop the date column for now - will add back in as a variable later
# Also Drop the RainWA column

siteDataClass = siteData[,-c(1,2,6,10)]
# get the number of observations/rows and multiply that by the proportion for train/test to split the data

# set the test proportion to 10%  
testProp = 0.10
# need to round so we get whole number  0 dp
trainDataLength = round(nrow(siteData)*(1-testProp),0)

train = siteDataClass[1:trainDataLength,]
test =  siteDataClass[(trainDataLength+1):nrow(siteData),]

```

###### Use the OTSC library to generate new positive samples
Note I have modified the functions from this package so they work with our data


```{r}
library(OSTSC)
library(pdist)
library(MASS)
library(fields)

#  Separate class variable from predictor variables ****

train.label = train$class
train.sample = as.matrix(train[,-7])
test.label = test$class
test.sample =  as.matrix(test[,-7])

dim(train.sample)

table(train.label)

# generate the new training samples

OversampledTrain = JillOSTSC(train.sample, train.label, parallel = FALSE, per=0.50)

(table(OversampledTrain[["label"]]))

# get the data into a dataframe for passing to the model

osTrain = as.data.frame(OversampledTrain, optional=FALSE)
names(osTrain) = names(train)

# Write the oversampled data to csv file

filenametrain = paste0("C:/Users/j/Documents/Uni/ComSci 760/Weather Project/Data/" ,
                          site, "_oversampledTrainingData.csv")

write.csv(osTrain,filenametrain, row.names = FALSE)

```

Have a look at the oversampled data

```{r}
# Histograms 

par(mfrow=c(3, 3))

for (i in 1:7) {
    hist(osTrain[,i], main= paste0(site," ",names(osTrain)[i]), col = 'blue')
}

# put back to 1 by 1 charts
par(mfrow=c(1, 1))
```

```{r}

# Density plot
par(mfrow=c(3, 3))

for (i in 1:7) {
    d =density(osTrain[,i], na.rm = TRUE)
    plot(d, type="n", main= paste0(site," ",names(osTrain)[i]))
    polygon(d, col="red", border="gray")
}
```

```{r}

######  build a multiple logistic regression model

# fig a glm to resp2
# fit the model
fit_multLR = glm(class ~ ., data=osTrain, family=binomial)
# get the predicted values
yhat_multLR= as.numeric(predict(fit_multLR, newdata=test, type = 'response')>0.5)
head(yhat_multLR)
# show model summary
summary(fit_multLR)
# show the confusion table
(confusionMatrix = table(test$class, yhat_multLR))

modelAccuracy = sum(diag(confusionMatrix))/sum(confusionMatrix)
# True positive rate  
modelSensitivity = confusionMatrix[2,2]/rowSums(confusionMatrix)[2]
# True negative rate  
modelSepicificy = confusionMatrix[1,1]/rowSums(confusionMatrix)[1]

cat("model accuracy = " ,modelAccuracy*100, "% ")
cat("model Sensitivity  = " ,modelSensitivity*100, "% ")
cat("model Sepicificy = " ,modelSepicificy*100, "% ") 
```
```{r}

######  build a multiple logistic regression model

# fig a glm to resp2
# fit the model
fit_multLR = glm(class ~ ., data=train, family=binomial)
# get the predicted values
yhat_multLR= as.numeric(predict(fit_multLR, newdata=test, type = 'response')>0.5)
head(yhat_multLR)
# show model summary
summary(fit_multLR)
# show the confusion table
(confusionMatrix = table(test$class, yhat_multLR))

(modelAccuracy = sum(diag(confusionMatrix))/sum(confusionMatrix))
# True positive rate  
(modelSensitivity = confusionMatrix[2,2]/rowSums(confusionMatrix)[2])
# True negative rate  
(modelSepicificy = confusionMatrix[1,1]/rowSums(confusionMatrix)[1])

cat("model accuracy = " ,modelAccuracy*100, "% ")
cat("model Sensitivity  = " ,modelSensitivity*100, "% ")
cat("model Sepicificy = " ,modelSepicificy*100, "% ") 

```


